### 2020-02-19 s1_text/s1_text.tex
KL edits to SI re reviewer comments

\newpage
\linenumbers

@@

Furthermore, we do not expect our approximation to have a large influence: the interaction energy will make the doubly bound state dominate, and the polymerase-XapR interaction should be weak in the singly bound state, because very weak activation is observed in the experiment when one of the XapR binding sites is removed~\cite{Chure2019}. If $e^{\Delta \epsilon_{\n{X1}}}$ and $e^{\Delta \epsilon_{\n{X2}}}$ are small, we have $w_{\n{p}} w_{\n{X1}} e^{\Delta \epsilon_{\n{X1}}} \approx w_{\n{p}} w_{\n{X2}} e^{\Delta \epsilon_{\n{X2}}} \approx w_{\n{p}} w_{\n{X}}$. If, in addition, the transcription rates for the singly bound states are much smaller than those for the doubly bound state, we can neglect the singly bound states in the nominator of $p_{\n{active}}$ (which, as argued in the previous paragraph ``Simplistic modeling of transcription'', is a good assumption). Thereby, we can obtain the same form of $p_{\n{active}}$ that we are working with in the main part by including the interaction between polymerase and XapR into $\epsilon_{\n{coop}}$.

@@

@@

\section{Additional plots and explanations of the results}
\paragraph*{The 2D set of equations.}
If the mRNA concentration is set to steady-state as discussed in the main text, the dynamical system is reduced to the two equations shown below:
\begin{alignat}{2}
\dd{\mathrm{[p]_a}}{\tau} &= \ &&
\rho 
\frac{
	\mathrm{[XapR]_{R,A}}^2 
	e^{- \Delta \epsilon_{\text{coop}}}
}{
	1 + 
	2 \mathrm{[XapR]_{R,A}} +
	\mathrm{[XapR]_{R,A}}^2 e^{- \Delta \epsilon_{\text{coop}}}
}
- \mathrm{[p]_a}
\\
\dd{\mathrm{[x]_a}}{\tau} &= \ && \left(k_{\n{\beta,i}} \frac{\n{[c]_a}}{K_{\n{\beta,i}} + \n{[c]_a}} - k_{\n{\beta,e}} \frac{\n{[x]_a}}{K_{\n{\beta,e}} + \n{[x]_a}} - k_{\n{\alpha}} \frac{\n{[x]_a}}{1 + \n{[x]_a}}\right) \mathrm{[p]_a} \\ & && + k_{\n{\eta}} \left(\mathrm{[c]_a} - \xi \mathrm{[x]_a}\right)
\nonumber \\
\text{with} & && \raggedleft [\n{XapR}]_{\mathrm{R,\n{A}}} = [\n{XapR}]_{\mathrm{R}} \frac{\left(1 + \frac{\mathrm{[x]_{\n{a}}}}{K_{\n{\chi A}}}\right)^2}{\left(1 + \frac{\mathrm{[x]_{\n{a}}}}{K_{\n{\chi A}}}\right)^2 + e^{\Delta \epsilon_{\n{x}}} \left(1+\frac{\mathrm{[x]_{\n{a}}}}{K_{\n{\chi A}}} \frac{1}{K_{\n{IA}}}\right)^2}. \nonumber
\end{alignat}
Here, several of the parameters from the 3D system could now be combined into one by defining $\rho \defeq \frac{\rho_{\n{m}} \rho_{\n{p}}}{\gamma_{\n{mp}}}$.

@@

@@

\FloatBarrier

@@

	\bibitem{Chure2019}
	Chure G, et~al.
	\newblock Measurements on the \emph{xapABR} system in \emph{E. coli.};
	Unpublished.

>>>>

\newpage

@@

Furthermore, we do not expect our approximation to have a large influence: the interaction energy will make the doubly bound state dominate, and the polymerase-XapR interaction should be weak in the singly bound state, because very weak activation is observed in the experiment when one of the XapR binding sites is removed (see section~\ref{sec:exp} in this supporting information text). If $e^{\Delta \epsilon_{\n{X1}}}$ and $e^{\Delta \epsilon_{\n{X2}}}$ are small, we have $w_{\n{p}} w_{\n{X1}} e^{\Delta \epsilon_{\n{X1}}} \approx w_{\n{p}} w_{\n{X2}} e^{\Delta \epsilon_{\n{X2}}} \approx w_{\n{p}} w_{\n{X}}$. If, in addition, the transcription rates for the singly bound states are much smaller than those for the doubly bound state, we can neglect the singly bound states in the nominator of $p_{\n{active}}$ (which, as argued in the previous paragraph ``Simplistic modeling of transcription'', is a good assumption). Thereby, we can obtain the same form of $p_{\n{active}}$ that we are working with in the main part by including the interaction between polymerase and XapR into $\epsilon_{\n{coop}}$.

@@

\FloatBarrier
\newpage

@@

\FloatBarrier
\newpage

\section{Additional plots and explanations of the results}
\paragraph*{The 2D set of equations.}
If the mRNA concentration is set to steady-state as discussed in the main text, the dynamical system is reduced to the two equations shown below:
\begin{alignat}{2}
\dd{\mathrm{[p]_a}}{\tau} &= \ &&
\rho 
\frac{
	\mathrm{[XapR]_{R,A}}^2 
	e^{- \Delta \epsilon_{\text{coop}}}
}{
	1 + 
	2 \mathrm{[XapR]_{R,A}} +
	\mathrm{[XapR]_{R,A}}^2 e^{- \Delta \epsilon_{\text{coop}}}
}
- \mathrm{[p]_a}
\\
\dd{\mathrm{[x]_a}}{\tau} &= \ && \left(k_{\n{\beta,i}} \frac{\n{[c]_a}}{K_{\n{\beta,i}} + \n{[c]_a}} - k_{\n{\beta,e}} \frac{\n{[x]_a}}{K_{\n{\beta,e}} + \n{[x]_a}} - k_{\n{\alpha}} \frac{\n{[x]_a}}{1 + \n{[x]_a}}\right) \mathrm{[p]_a} \\ & && + k_{\n{\eta}} \left(\mathrm{[c]_a} - \xi \mathrm{[x]_a}\right)
\nonumber \\
\text{with} & && \raggedleft [\n{XapR}]_{\mathrm{R,\n{A}}} = [\n{XapR}]_{\mathrm{R}} \frac{\left(1 + \frac{\mathrm{[x]_{\n{a}}}}{K_{\n{\chi A}}}\right)^2}{\left(1 + \frac{\mathrm{[x]_{\n{a}}}}{K_{\n{\chi A}}}\right)^2 + e^{\Delta \epsilon_{\n{x}}} \left(1+\frac{\mathrm{[x]_{\n{a}}}}{K_{\n{\chi A}}} \frac{1}{K_{\n{IA}}}\right)^2}. \nonumber
\end{alignat}
Here, several of the parameters from the 3D system could now be combined into one by defining $\rho \defeq \frac{\rho_{\n{m}} \rho_{\n{p}}}{\gamma_{\n{mp}}}$.
\FloatBarrier

@@

\paragraph*{Including transcriptional bursting.} To test the robustness of our results to larger stochasticity (which is expected in the real biological system), we included some burstiness in transcription in our stochastic simulations. We assumed a geometric distribution for the burst size $s$, i.e. $p(s) = (1/4)\cdot (1-1/4)^{s-1}$ with mean $4$, which is a bit larger than what usually occurs in biology \textcolor{red}{citation missing}. We then cut this distribution at a burst size of 10, meaning we did not consider larger bursts, leading to an actual mean of around $3.2$. The transcription rate was rescaled such that the mean number of mRNA's produced remained the same.

We observe the same qualitative behavior but with stronger fluctuations around the mean and also larger variation in the time to the steady-state mean. As an example, Fig~\ref{evolbursts} shows the time evolution of two runs of the simulation with bursting for the same conditions (i.e. parameters and starting values) as the time evolution that was shown in the main text. Fig~\ref{distbursts} shows the distribution obtained for the same and an additional condition (i.e. parameters and starting values) as for the distributions shown in the main text. The variance in the timescales and distributions is larger than without bursting. Furthermore, switching already happens for lower $\mathrm{[c]_a}$: in \subref{distbursts}{C}, most cells are in the high expression state, whereas in the corresponding distribution in the main text, there were many uninduced cells, i.e. the first peak was higher. To obtain a bimodal distribution like in the main text, $\mathrm{[c]_a}$ now needs to be set to $17.5$ instead of $18.5$ as shown in \subref{distbursts}{B}. This was to be expected, since the larger fluctuations make it easier for the system to jump across the middle fixed point.

Still, there are no big qualitative differences to the results without bursting. Thus, the assumptions we made that lead to less stochasticity than in the real biological system are not expected to have a very significant influence on our results, and the switch-like behavior seems to be stable to stochastic perturbations. \textcolor{red}{Fig 7 C not enough runs yet}

\begin{figure}
	\centering
	\includegraphics[width=1\textwidth]{FigSI6_7.pdf}
	\caption{{\bf Time evolution from the stochastic simulation when transcriptional bursting is included.}
		(A) and (B) each show the time evolution from one run of the simulation. The simulation was run once for $5 \cdot 10^5 \unit{s}$ and started
		at an mRNA, protein and intracellular xanthosine count of 0. The parameters that were used are the same as in the
		table in the main text, the only exception being the
		extracellular xanthosine concentration, which was chosen to be
		$\n{[c]_a} = 25$ (recall $\n{[c]_{a}} \defeq \frac{c}{K_{\n{a}}}$,
		$K_{\n{a}} = 5 \cdot 10^{-5} \unit{M}$).}
	\label{evolbursts}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=1\textwidth]{FigSI6_7_2.pdf}
	\caption{{\bf Distributions from the stochastic simulation when transcriptional bursting is included.}
		(A) shows the distribution obtained from the simulation with $\n{[c]_a}=12$, (B), (C), and (D) show the same for $\n{[c]_a}=17.5$, $\mathrm{[c]_a}=18.5$, and $\mathrm{[c]_a}=25$, respectively (recall $\n{[c]_{a}} \defeq \frac{c}{K_{\n{a}}}$,
		$K_{\n{a}} = 5 \cdot 10^{-5} \unit{M}$). All other parameters are as in the table in the main text, and the simulation
		was run 1000 times for $10^6 \unit{s}$ each (simulated time) and
		started at a mRNA, protein and intracellular xanthosine count of 0.}
	\label{distbursts}
\end{figure}

@@

\FloatBarrier
\newpage

\section{Experimental materials and methods} \label{sec:exp}

\textcolor{red}{missing}

\FloatBarrier
\newpage

@@
